---
title: "Bar Chart Race for Refugees"
author: "Edouard Legoupil"
date: "2019-06-08"

  
---



```{r setup, include = FALSE, echo = FALSE, warning = FALSE}
using <- function(...) {
    libs <- unlist(list(...))
    req <- unlist(lapply(libs,require,character.only = TRUE))
    need <- libs[req == FALSE]
    if (length(need) > 0) { 
        install.packages(need)
        lapply(need,require,character.only = TRUE)
    }
}

using('tidyverse','gganimate','gghighlight','ggpubr', 'dplyr', 'tidyr', 'gapminder', 'ggplot2',  'ggalt', 'forcats', 'R.utils', 'png', 'grid', 'ggpubr', 'scales', 'bbplot', 'markdown', 'pander', 'ISOcodes', 'wbstats', 'sf', 'rnaturalearth', 'rnaturalearthdata', 'ggspatial')

options(scipen = 999) # turn-off scientific notation like 1e+48
```



``` {r , echo = FALSE}
unhcr_style <- function() {
  font <- "Lato"
  ggplot2::theme(
    
#This sets the font, size, type and colour of text for the chart's title
  plot.title = ggplot2::element_text(family=font, size=20, face = "bold", color = "#222222"),

#This sets the font, size, type and colour of text for the chart's subtitle,  as well as setting a margin between the title and the subtitle
  plot.subtitle = ggplot2::element_text(family=font, size=16, margin=ggplot2::margin(9,0,9,0)),
  plot.caption = ggplot2::element_blank(),

#This sets the position and alignment of the legend, removes a title and backround for it and sets the requirements for any text within the legend. The legend may often need some more manual tweaking when it comes to its exact position based on the plot coordinates.
  legend.position = "top",
  legend.text.align = 0,
  legend.background = ggplot2::element_blank(),
  legend.title = ggplot2::element_blank(),
  legend.key = ggplot2::element_blank(),
  legend.text = ggplot2::element_text(family=font, size=13, color = "#222222"),

#This sets the text font, size and colour for the axis test, as well as setting the margins and removes lines and ticks. In some cases, axis lines and axis ticks are things we would want to have in the chart
  axis.title = ggplot2::element_blank(),
  axis.text = ggplot2::element_text(family=font, size=13, color = "#222222"),
  axis.text.x = ggplot2::element_text(margin=ggplot2::margin(5, b = 10)),
  axis.ticks = ggplot2::element_blank(),
  axis.line = ggplot2::element_blank(),

#This removes all minor gridlines and adds major y gridlines. In many cases you will want to change this to remove y gridlines and add x gridlines. 
  panel.grid.minor = ggplot2::element_blank(),
  panel.grid.major.y = ggplot2::element_line(color = "#cbcbcb"),
  panel.grid.major.x = ggplot2::element_blank(),

#This sets the panel background as blank, removing the standard grey ggplot background colour from the plot
  panel.background = ggplot2::element_blank(),

#This sets the panel background for facet-wrapped plots to white, removing the standard grey ggplot background colour and sets the title size of the facet-wrap title to font size 22
  strip.background = ggplot2::element_rect(fill = "white"),
  strip.text = ggplot2::element_text(size  = 13,  hjust = 0)
  )
}

#Left align text
left_align <- function(plot_name, pieces){
  grob <- ggplot2::ggplotGrob(plot_name)
  n <- length(pieces)
  grob$layout$l[grob$layout$name %in% pieces] <- 2
  return(grob)
}

## a little help function to better format numbers
format_si <- function(...) {
  function(x) {
    limits <- c(1e-24, 1e-21, 1e-18, 1e-15, 1e-12,
                1e-9,  1e-6,  1e-3,  1e0,   1e3,
                1e6,   1e9,   1e12,  1e15,  1e18,
                1e21,  1e24)
    prefix <- c("y",   "z",   "a",   "f",   "p",
                "n",   "",   "m",   " ",   "k",
                "M",   "G",   "T",   "P",   "E",
                "Z",   "Y")

    # Vector with array indices according to position in intervals
    i <- findInterval(abs(x), limits)

    # Set prefix to " " for very small values < 1e-24
    i <- ifelse(i == 0, which(limits == 1e0), i)

    paste(format(round(x/limits[i], 1),
                 trim = TRUE, scientific = FALSE, ...),
          prefix[i])
  }
}
```




"Bar Chart Race" are a specific type of bar chart that moves to show rankings over time. It became recently a quite popular approach to bring a storytelling elements within a longitudinal dataset. Readers are suggested to connect and combine what they see on the chart with other qualitive elements that they know about (elements of history). By using the allegory of  F1 Race, it gives a very dynamic dimension. 




With R, it's fairly easy to reproduce such vis - We will show here how to use `gganimate` R packages to display evolution of refugees & IDP's number over time based on UNHCR statistics




The visualisation below uses data from UNHCR popstat and data.worldbank API

```{r getdata, message=FALSE, warning=FALSE, echo = FALSE}
# Time series
#url <- paste( 'http://popstats.unhcr.org/en/time_series.csv') 
#download.file(url, destfile = "unhcr_popstats_export_time_series_all_data.csv" )



## Get region name
#url <- paste( 'https://pkgstore.datahub.io/core/country-codes/country-codes_csv/data/3b9fd39bdadd7edd7f7dcee708f47e1b/country-codes_csv.csv') 
#download.file(url, destfile = "countrycode.csv" )

# # Population, GDP & GNP per Capita from WorldBank
# wb_data <- wb( indicator = c("SP.POP.TOTL", "NY.GDP.MKTP.CD", "NY.GDP.PCAP.CD", "NY.GNP.PCAP.CD"),
#                startdate = 1951, enddate = 2017, return_wide = TRUE)
# # Renaming variables for further matching
# names(wb_data)[1] <- "ISO3166.1.Alpha.3"
# names(wb_data)[2] <- "Year"
# 
# write.csv(wb_data, "wb_data.csv", row.names = FALSE)

```

Reshape the data to get the top 10 for each year using tidyverse

```{r reshapedata, message=FALSE, warning=FALSE, echo = FALSE}

time_series <- read.csv("unhcr_popstats_export_time_series_all_data.csv", skip = 3) 
## rename the country and make sure Value is a number...
names(time_series)[2] <- "Country"

## make sure Value is numeric
time_series$Value <- as.integer(as.character(time_series$Value))
time_series$Population.type <- as.factor(as.character(time_series$Population.type))


wb_data <- read.csv("wb_data.csv") 
countrycode <- read.csv("countrycode.csv") 
#names(countrycode)

### Rewrite country name in UNHCR dataset for further matching
time_series$ctryiso <- as.character(time_series$Country)
time_series$ctryiso[time_series$Country == "Bonaire"] <- "Bonaire, Sint Eustatius and Saba" 
time_series$ctryiso[time_series$Country == "Central African Rep."] <- "Central African Republic" 
time_series$ctryiso[time_series$Country == "China, Hong Kong SAR"] <- "China, Hong Kong Special Administrative Region" 
time_series$ctryiso[time_series$Country == "China, Macao SAR"] <- "China, Macao Special Administrative Region"  
time_series$ctryiso[time_series$Country == "C\xf4te d'Ivoire"] <- "Côte d'Ivoire" 
time_series$ctryiso[time_series$Country == "Cura\xe7ao"] <- "Curaçao" 
time_series$ctryiso[time_series$Country == "Czech Rep."] <- "Czechia" 
time_series$ctryiso[time_series$Country == "Dem. Rep. of the Congo"] <- "Democratic Republic of the Congo" 
time_series$ctryiso[time_series$Country == "Dominican Rep."] <- "Dominican Republic" 
time_series$ctryiso[time_series$Country == "Iran (Islamic Rep. of)"] <- "Iran (Islamic Republic of)" 
time_series$ctryiso[time_series$Country == "Lao People's Dem. Rep."] <- "Lao People's Democratic Republic" 
time_series$ctryiso[time_series$Country == "Rep. of Korea"] <- "Republic of Korea" 
time_series$ctryiso[time_series$Country == "Dem. People's Rep. of Korea"] <- "Democratic People's Republic of Korea" 
time_series$ctryiso[time_series$Country == "Rep. of Moldova"] <- "Republic of Moldova"  
time_series$ctryiso[time_series$Country == "Serbia and Kosovo (S/RES/1244 (1999))"] <- "Serbia" 
time_series$ctryiso[time_series$Country == "Syrian Arab Rep."] <- "Syrian Arab Republic" 
time_series$ctryiso[time_series$Country == "United Rep. of Tanzania"] <- "United Republic of Tanzania" 
time_series$ctryiso[time_series$Country == "Holy See (the)"] <- "Holy See (Vatican City State)" 
time_series$ctryiso[time_series$Country == "Réunion"] <- "Reunion" 
time_series$ctryiso[time_series$Country == "Saint-Pierre-et-Miquelon"] <- "Saint Pierre and Miquelon" 
time_series$ctryiso[time_series$Country == "US Virgin Islands"] <- "Virgin Islands, U.S." 
time_series$ctryiso[time_series$Country == "Wallis and Futuna Islands "] <- "Wallis and Futuna" 
time_series$ctryiso[time_series$Country == "United Kingdom"] <- "United Kingdom of Great Britain and Northern Ireland"

time_series <- merge(x = time_series , by.x = "ctryiso", all.x = TRUE, y = countrycode , by.y = "official_name_en" )

## Check what population type we have there and subset accordingly
#levels(time_series$Population.type)

time_series2 <- time_series[ time_series$Population.type %in% c("Refugees (incl. refugee-like situations)", "Internally displaced persons" ), ]

time_series2$Country <- as.character(time_series2$Country)
time_series2$Country[time_series2$Country == "C\xf4te d'Ivoire"] <- "Cote d'Ivoire" 
time_series2$Country <- as.factor(time_series2$Country)

## Remove unknow countries or blanks
time_series2 <- time_series2[ !(time_series2$Country %in% c("Various/Unknown" )), ]

## Remove if value is 0 or NA
time_series2 <- time_series2[ !(is.na(time_series2$Value)), ]
time_series2 <- time_series2[ time_series2$Value != 0, c("ctryiso", "Year", "Country",  
                                                         "Origin", "Population.type", "Value",
                                                         "ISO3166.1.Alpha.3", "Developed...Developing.Countries",
                                                         "Region.Name", "Sub.region.Name"   )  ]

#names(time_series2)


## Now get the rank from high to low for all countries per  year and population type
rank_data <- time_series2 %>%
  group_by(Year, Population.type, Country, ISO3166.1.Alpha.3, Region.Name) %>%
  summarise(Value2 = sum(Value) ) 

#str(as.data.frame(rank_data))
rank_data <- as.data.frame(rank_data)


rank_data1 <- merge(x = rank_data, y = wb_data, by = c("ISO3166.1.Alpha.3", "Year"), all.x = TRUE)
rank_data1$Ref.per.pop <- round(rank_data1$Value2 / rank_data1$SP.POP.TOTL *100 , 2)
rank_data1$Ref.per.gdp <- round(rank_data1$NY.GDP.MKTP.CD / rank_data1$Value2 , 2)
rank_data1$Ref.per.gdp.cap <-  round(rank_data1$NY.GDP.PCAP.CD / rank_data1$Value2 , 2)



rank_data.ref <- rank_data1 %>%
  group_by(Year, Population.type) %>%
  
  ## Tried first rank but did not provided ranks as integer... 
  # mutate(ordering = rank(-Value), ties.method = "min") %>%
  
  mutate(ordering = rank(-Value2)) %>%
  ungroup() 

## check our value for rank -- Note that there are different tie method
#levels(as.factor(rank_data$ordering))


rank_data.ref$Country <- as.character(rank_data.ref$Country)


## Need to fix manually issue when ex-aequo rank rank = 8.5 

## In 1962
rank_data.ref$ordering[rank_data.ref$ordering == 10 &
                      rank_data.ref$Year == "1962" &
                      rank_data.ref$Population.type %in% c("Refugees (incl. refugee-like situations)")] <- 11

rank_data.ref$ordering[rank_data.ref$ordering == 8.5 &
                      rank_data.ref$Year == "1962" &
                      rank_data.ref$Country == "Burundi" &
                      rank_data.ref$Population.type %in% c("Refugees (incl. refugee-like situations)")] <- 9

rank_data.ref$ordering[rank_data.ref$ordering == 8.5 &
                      rank_data.ref$Year == "1962" &
                      rank_data.ref$Country == "Austria" &
                      rank_data.ref$Population.type %in% c("Refugees (incl. refugee-like situations)")] <- 10


## In 1978
rank_data.ref$ordering[rank_data.ref$ordering == 10 &
                      rank_data.ref$Year == "1978" &
                      rank_data.ref$Population.type %in% c("Refugees (incl. refugee-like situations)")] <- 11

rank_data.ref$ordering[rank_data.ref$ordering == 8.5 &
                      rank_data.ref$Year == "1978" &
                      rank_data.ref$Country == "Viet Nam" &
                      rank_data.ref$Population.type %in% c("Refugees (incl. refugee-like situations)")] <- 9

rank_data.ref$ordering[rank_data.ref$ordering == 8.5 &
                      rank_data.ref$Year == "1978" &
                      rank_data.ref$Country == "United Kingdom" & 
                      rank_data.ref$Population.type %in% c("Refugees (incl. refugee-like situations)")] <- 10


## and for IPDs


## In 1996
rank_data.ref$ordering[rank_data.ref$ordering == 10 &
                      rank_data.ref$Year == "1996" &
                      rank_data.ref$Population.type %in% c("Internally displaced persons")] <- 11

rank_data.ref$ordering[rank_data.ref$ordering == 9.5 &
                      rank_data.ref$Year == "1996" &
                      rank_data.ref$Country == "Somalia" &
                      rank_data.ref$Population.type %in% c("Internally displaced persons")] <- 11


rank_data.ref$Country[rank_data.ref$ordering == 9.5 &
                      rank_data.ref$Year == "1996" &
                      rank_data.ref$Country == "Sri Lanka" & 
                      rank_data.ref$Population.type %in% c("Internally displaced persons")] <- "Sri Lanka / Somalia"

rank_data.ref$ordering[rank_data.ref$ordering == 9.5 &
                      rank_data.ref$Year == "1996" &
                      rank_data.ref$Country == "Sri Lanka / Somalia" & 
                      rank_data.ref$Population.type %in% c("Internally displaced persons")] <- 10

## in 1997
rank_data.ref$ordering[rank_data.ref$ordering == 10 &
                      rank_data.ref$Year == "1997" &
                      rank_data.ref$Population.type %in% c("Internally displaced persons")] <- 11

rank_data.ref$ordering[rank_data.ref$ordering == 9.5 &
                      rank_data.ref$Year == "1997" &
                      rank_data.ref$Country == "Somalia" &
                      rank_data.ref$Population.type %in% c("Internally displaced persons")] <- 11


rank_data.ref$Country[rank_data.ref$ordering == 9.5 &
                      rank_data.ref$Year == "1997" &
                      rank_data.ref$Country == "Sri Lanka" & 
                      rank_data.ref$Population.type %in% c("Internally displaced persons")] <- "Sri Lanka / Somalia"

rank_data.ref$ordering[rank_data.ref$ordering == 9.5 &
                      rank_data.ref$Year == "1997" &
                      rank_data.ref$Country == "Sri Lanka / Somalia" & 
                      rank_data.ref$Population.type %in% c("Internally displaced persons")] <- 10


#rank_data.ref$Country <- as.factor(rank_data.ref$Country)
## Regnerate factors modality - 
rank_data.ref$Country <- as.factor(as.character(rank_data.ref$Country))

# Filter only top 10 
rank_data.ref <- rank_data.ref[rank_data.ref$ordering <= 10, ]
#rank_data$Year = as.Date(as.character(rank_data$Year), format = "%Y")


# levels(as.factor(rank_data.ref$Country))
## Double checking country name
#table(time_series2$Country, useNA = "ifany")

## rank_data1$Ref.per.pop rank_data1$Ref.per.gdp rank_data1$Ref.per.gdp.cap

#levels(rank_data.ref$Region.Name)

rank_data.ref$Region.Name <- as.factor(as.character(rank_data.ref$Region.Name))

#levels(rank_data.ref$Region.Name)

```

# Displaced Population Race

## In 1951

```{r chart1951, message=FALSE, warning=FALSE, echo = FALSE}


ggplot(rank_data.ref[ rank_data.ref$Year == 1951 & rank_data.ref$Population.type %in% c("Refugees (incl. refugee-like situations)"), ]) +
  geom_bar(aes(y = Value2,  x =   reorder(ordering, desc(ordering)),
               group = Country ,color = Country, fill = Country), alpha = 0.75, stat = "identity") +
  
  geom_label(aes(y = 0 , x =  reorder(ordering, desc(ordering)), label = Country),
             hjust = 0,
             vjust = 0.5,
             fill = NA,
             label.size = NA,
             family = "Helvetica",
             size = 6) +

  coord_flip(clip = "off", expand = FALSE) +
  scale_color_viridis_d(option = "plasma" ) +
  scale_fill_viridis_d(option = "plasma") +
  scale_y_continuous(labels = format_si()) +
  theme_minimal(14, "Avenir") +
  
  guides(color = F, fill = F) +
  labs(title =  "Top 10 Hosting Countries",
       subtitle = 'Year 1951',
       y = "Population Size",
       x = "",
       caption =  "Source: UNHCR Population Statistics -http://popstats.unhcr.org ") +
  theme(plot.title = element_text(hjust = 1, size = 22),
        axis.ticks.y = element_blank(),
        axis.text.y  = element_blank(), 
        panel.background  = element_blank(), 
        panel.grid = element_blank(),
        plot.background = element_blank(),
        legend.position = "bottom",
        panel.grid.major.x = element_line(color = "#cbcbcb"), 
        panel.grid.major.y = element_blank()) 

```

## by Population Group for 2017

```{r chart2017facet, message=FALSE, warning=FALSE, echo = FALSE}


ggplot(rank_data.ref[ rank_data.ref$Year == 2017, ]) +
  geom_bar(aes(y = Value2,  x =   reorder(ordering, desc(ordering)),
               group = Country ,color = Country, fill = Country), alpha = 0.75, stat = "identity") +
  
  geom_label(aes(y = 0 , x =  reorder(ordering, desc(ordering)), label = Country),
             hjust = 0,
             vjust = 0.5,
             fill = NA,
             label.size = NA,
             family = "Helvetica",
             size = 6) +
  
  #facet_wrap( ~ Population.type) +
  facet_grid(. ~ Population.type) +
  coord_flip(clip = "off", expand = FALSE) +
  scale_color_viridis_d(option = "plasma" ) +
  scale_fill_viridis_d(option = "plasma") +
  scale_y_continuous(labels = format_si()) +
  theme_minimal(14, "Avenir") +
  
  guides(color = F, fill = F) +
  labs(title =  "Top 10 Countries",
       subtitle = 'Year 2017',
       y = "Population Size",
       x = "",
       caption =  "Source: UNHCR Population Statistics -http://popstats.unhcr.org ") +
  theme(plot.title = element_text(hjust = 1, size = 22),
        axis.ticks.y = element_blank(),
        axis.text.y  = element_blank(), 
        panel.background  = element_blank(), 
        panel.grid = element_blank(),
        plot.background = element_blank(),
        legend.position = "bottom",
        panel.grid.major.x = element_line(color = "#cbcbcb"), 
        panel.grid.major.y = element_blank()) 

```




## Animate

For animation, the main challenge is to tune parameters. 


 * use geom_tile, not geom_bar as this allows for better transitions within gganimate
 * gganimate functions transition_time and ease_aes handle the animation and transition between bars. The settings here worked well for my purposes, but dig into these functions to get an overview of different options
 *    nframes and fps parameters to the animate function control the speed of transitions. One mistake I made here initially was to set nframes equal to the number of years in the dataset. This works, but because there is only 1 frame per year, you don't get the smooth transitions that I wanted in this graph. Increasing the number of frames fixed that issue.

nframes	: The number of frames to render (default 100)

fps	: The framerate of the animation in frames/sec (default 10)

and animate.... the bar chart race... 

```{r barchartrace , message=FALSE, warning=FALSE, echo = FALSE}

p <- rank_data.ref[rank_data.ref$Population.type %in% c("Refugees (incl. refugee-like situations)"), ] %>%
  ggplot(aes(y = Value2,  x =   reorder(ordering, desc(ordering)), group = Country ,)) +
  geom_tile(aes(y = Value2 / 2, height = Value2, fill = Region.Name), width = 0.9) +
  geom_text(aes(label = Country), hjust = "right", colour = "black", fontface = "bold", nudge_y = -100000) +
  geom_text(aes(label = scales::comma(Value2)), hjust = "left", nudge_y = 100000, colour = "grey30") +
  coord_flip(clip = "off") +
  scale_fill_manual(name = 'Region.Name', values = c("#7fc97f", "#beaed4", "#fdc086", "#ffff99", "#386cb0")) +
  scale_x_discrete("") +
  scale_y_continuous("",labels = format_si()) +
  hrbrthemes::theme_ipsum(plot_title_size = 32, subtitle_size = 24, caption_size = 12, base_size = 22) +
  theme(panel.grid.major.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        legend.position = c(0.8, 0.2),
        legend.title = element_blank(),
        plot.margin = margin(1,1,1,3,"cm"),
        axis.text.y = element_blank()) +
  # gganimate code to transition by year:
  transition_time(Year) +
  ease_aes('cubic-in-out') +
  labs(title = 'Top 10 Refugee Hosting Countries',
       subtitle = 'Refugee Population in {round(frame_time,0)}',
       caption = 'Source: UNHCR http://popstats.unhcr.org
Visualisation:  @edouard_lgp')

animate(p, nframes = 750, fps = 20, end_pause = 50, width = 1000)
#anim_save("bar_chart_race_refugee.gif", animate(p, nframes = 750, fps = 20, end_pause = 60, width = 1000))

# save as preferred rendered format
#gganimate::animate(p, nframes = 200, fps = 5, duration = 220, width = 2000, height = 1200, renderer = ffmpeg_renderer("anim.mp4"))

```

## Animate IDPs..

```{r barchartrace2 , message=FALSE, warning=FALSE, echo = FALSE}

p <- rank_data.ref[rank_data.ref$Population.type %in% c("Internally displaced persons"), ] %>%
  ggplot(aes(y = Value2,  x =   reorder(ordering, desc(ordering)), group = Country ,)) +
  geom_tile(aes(y = Value2 / 2, height = Value2, fill = Region.Name), width = 0.9) +
  geom_text(aes(label = Country), hjust = "right", colour = "black", fontface = "bold", nudge_y = -100000) +
  geom_text(aes(label = scales::comma(Value2)), hjust = "left", nudge_y = 100000, colour = "grey30") +
  coord_flip(clip = "off") +
  scale_fill_manual(name = 'Region.Name', values = c("#7fc97f", "#beaed4", "#fdc086", "#ffff99", "#386cb0")) +
  scale_x_discrete("") +
  scale_y_continuous("",labels = format_si()) +
  hrbrthemes::theme_ipsum(plot_title_size = 32, subtitle_size = 24, caption_size = 12, base_size = 22) +
  theme(panel.grid.major.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        legend.position = c(0.8, 0.2),
        legend.title = element_blank(),
        plot.margin = margin(1,1,1,3,"cm"),
        axis.text.y = element_blank()) +
  # gganimate code to transition by year:
  transition_time(Year) +
  ease_aes('cubic-in-out') +
  labs(title = 'Top 10 Countries hosting Internally Displaced Persons',
       subtitle = 'Displaced Population in {round(frame_time,0)}',
       caption = 'Source: UNHCR http://popstats.unhcr.org
Visualisation:  @edouard_lgp')

animate(p, nframes = 750, fps = 20, end_pause = 50, width = 1000)

```


# Race Weighted by Total population


```{r preparedata2, message=FALSE, warning=FALSE, echo = FALSE}


rank_data.pop <- rank_data1 %>%
    filter( !(is.na(Ref.per.pop))) %>%
    group_by(Year, Population.type) %>%
    mutate(ordering = rank(-Ref.per.pop)) %>%
    ungroup() %>%
    select(Country, Year, Ref.per.pop, ordering, Value2, Population.type, Region.Name)

rank_data.pop$Country <- as.factor(as.character(rank_data.pop$Country))
rank_data.pop <- rank_data.pop[rank_data.pop$ordering <= 10, ]

rank_data.pop$Region.Name <- as.factor(as.character(rank_data.pop$Region.Name))
#rank_data.pop$ordering <- as.integer(rank_data.pop$ordering)


rank_data.pop$Country <- as.character(rank_data.pop$Country)

# levels(as.factor(rank_data.pop$ordering))
# View(rank_data.pop[ rank_data.pop$ordering == "9.5" &
#                       rank_data.pop$Population.type %in% c("Refugees (incl. refugee-like situations)"), ])
# 
# View(rank_data.pop[  rank_data.pop$Year == "1972"&
#                        rank_data.pop$Population.type %in% c("Refugees (incl. refugee-like situations)"), ])

## Need to fix manually issue when ex-aequo rank 


## In 1964
rank_data.pop$ordering[rank_data.pop$ordering == 9.5 &
                         rank_data.pop$Year == "1964" &
                         rank_data.pop$Country == "Austria" &
                         rank_data.pop$Population.type %in% c("Refugees (incl. refugee-like situations)")] <- 9

rank_data.pop$ordering[rank_data.pop$ordering == 9.5 &
                         rank_data.pop$Year == "1964" &
                         rank_data.pop$Country == "Switzerland" &
                         rank_data.pop$Population.type %in% c("Refugees (incl. refugee-like situations)")] <- 10


## In 1967
rank_data.pop$ordering[rank_data.pop$ordering == 9.5 &
                         rank_data.pop$Year == "1967" &
                         rank_data.pop$Country == "France" &
                         rank_data.pop$Population.type %in% c("Refugees (incl. refugee-like situations)")] <- 8

rank_data.pop$ordering[rank_data.pop$ordering == 9.5 &
                         rank_data.pop$Year == "1967" &
                         rank_data.pop$Country == "Belgium" &
                         rank_data.pop$Population.type %in% c("Refugees (incl. refugee-like situations)")] <- 9

rank_data.pop$ordering[rank_data.pop$ordering == 9.5 &
                         rank_data.pop$Year == "1967" &
                         rank_data.pop$Country == "Sudan" &
                         rank_data.pop$Population.type %in% c("Refugees (incl. refugee-like situations)")] <- 10

rank_data.pop[rank_data.pop$ordering == 9.5 &
                         rank_data.pop$Year == "1967" &
                         rank_data.pop$Country == "	Switzerland" &
                         rank_data.pop$Population.type %in% c("Refugees (incl. refugee-like situations)")] <- NULL

## In 1972
rank_data.pop$ordering[rank_data.pop$ordering == 9.5 &
                         rank_data.pop$Year == "1972" &
                         rank_data.pop$Country == "Switzerland" &
                         rank_data.pop$Population.type %in% c("Refugees (incl. refugee-like situations)")] <- 9

rank_data.pop$ordering[rank_data.pop$ordering == 9.5 &
                         rank_data.pop$Year == "1972" &
                         rank_data.pop$Country == "	Zambia" &
                         rank_data.pop$Population.type %in% c("Refugees (incl. refugee-like situations)")] <- 10



## In 1973
rank_data.pop$ordering[rank_data.pop$ordering == 6.5 &
                         rank_data.pop$Year == "1973" &
                         rank_data.pop$Country == "Zambia" &
                         rank_data.pop$Population.type %in% c("Refugees (incl. refugee-like situations)")] <- 6

rank_data.pop$ordering[rank_data.pop$ordering == 6.5 &
                         rank_data.pop$Year == "1973" &
                         rank_data.pop$Country == "United Arab Emirates" &
                         rank_data.pop$Population.type %in% c("Refugees (incl. refugee-like situations)")] <- 7


## In 2003
rank_data.pop$ordering[rank_data.pop$ordering == 5.5 &
                         rank_data.pop$Year == "2003" &
                         rank_data.pop$Country == "Guinea" &
                         rank_data.pop$Population.type %in% c("Refugees (incl. refugee-like situations)")] <- 6

rank_data.pop$ordering[rank_data.pop$ordering == 5.5 &
                         rank_data.pop$Year == "2003" &
                         rank_data.pop$Country == "Zambia" &
                         rank_data.pop$Population.type %in% c("Refugees (incl. refugee-like situations)")] <- 5


## In 2009
rank_data.pop$ordering[rank_data.pop$ordering == 6.5 &
                         rank_data.pop$Year == "2009" &
                         rank_data.pop$Country == "Iran (Islamic Rep. of)" &
                         rank_data.pop$Population.type %in% c("Refugees (incl. refugee-like situations)")] <- 6

rank_data.pop$ordering[rank_data.pop$ordering == 6.5 &
                         rank_data.pop$Year == "2009" &
                         rank_data.pop$Country == "Djibouti" &
                         rank_data.pop$Population.type %in% c("Refugees (incl. refugee-like situations)")] <- 7

## Regnerate factors modality - 
rank_data.pop$Country <- as.factor(as.character(rank_data.pop$Country))

```


```{r chart2017facet2, message=FALSE, warning=FALSE, echo = FALSE}

ggplot(rank_data.pop[ rank_data.pop$Year == 2017, ]) +
  geom_bar(aes(y = Ref.per.pop,  x =   reorder(ordering, desc(ordering)),
               group = Country ,color = Country, fill = Country), alpha = 0.75, stat = "identity") +
  
  geom_label(aes(y = 0 , x =  reorder(ordering, desc(ordering)), label = Country),
             hjust = 0,
             vjust = 0.5,
             fill = NA,
             label.size = NA,
             family = "Helvetica",
             size = 6) +
  
  #facet_wrap( ~ Population.type) +
    facet_grid(. ~ Population.type) +
  coord_flip(clip = "off", expand = FALSE) +
  scale_color_viridis_d(option = "plasma" ) +
  scale_fill_viridis_d(option = "plasma") +
  scale_y_continuous(labels = format_si()) +
  theme_minimal(14, "Avenir") +
  
  guides(color = F, fill = F) +
  labs(title =  "Top 10 Countries Displaced Vs Population",
       subtitle = 'Year 2017',
       y = "Refugee to Population Ratio",
       x = "",
       caption =  "Source: UNHCR Population Statistics -http://popstats.unhcr.org - World Bank ") +
  theme(plot.title = element_text(hjust = 1, size = 22),
        axis.ticks.y = element_blank(),
        axis.text.y  = element_blank(), 
        panel.background  = element_blank(), 
        panel.grid = element_blank(),
        plot.background = element_blank(),
        legend.position = "bottom",
        panel.grid.major.x = element_line(color = "#cbcbcb"), 
        panel.grid.major.y = element_blank()) 

```


```{r chart2017pop, message=FALSE, warning=FALSE, echo = FALSE}

ggplot(rank_data.pop[ rank_data.pop$Year == 2017 & 
                        rank_data.pop$Population.type %in% c("Refugees (incl. refugee-like situations)"), ]) +
  geom_bar(aes(y = Ref.per.pop,  x =   reorder(ordering, desc(ordering)),
               group = Country ,color = Country, fill = Country), alpha = 0.75, stat = "identity") +
  
  geom_label(aes(y = 0 , x =  reorder(ordering, desc(ordering)), label = Country),
             hjust = 0,
             vjust = 0.5,
             fill = NA,
             label.size = NA,
             family = "Helvetica",
             size = 6) +
  
  #facet_wrap( ~ Population.type) +
  #facet_grid(. ~ Population.type) +
  coord_flip(clip = "off", expand = FALSE) +
  scale_color_viridis_d(option = "plasma" ) +
  scale_fill_viridis_d(option = "plasma") +
  scale_y_continuous(labels = format_si()) +
  theme_minimal(14, "Avenir") +
  
  guides(color = F, fill = F) +
  labs(title =  "Top 10 Countries Refugee per Population",
       subtitle = 'Year 2017',
       y = "Refugee to Population Ratio",
       x = "",
       caption =  "Source: UNHCR Population Statistics -http://popstats.unhcr.org - World Bank ") +
  theme(plot.title = element_text(hjust = 1, size = 22),
        axis.ticks.y = element_blank(),
        axis.text.y  = element_blank(), 
        panel.background  = element_blank(), 
        panel.grid = element_blank(),
        plot.background = element_blank(),
        legend.position = "bottom",
        panel.grid.major.x = element_line(color = "#cbcbcb"), 
        panel.grid.major.y = element_blank()) 

```



```{r barchartrace3 , message=FALSE, warning=FALSE, echo = FALSE}


p <- rank_data.pop[rank_data.pop$Population.type %in% c("Refugees (incl. refugee-like situations)"), ] %>%
  ggplot(aes(y = Ref.per.pop,  x =   reorder(ordering, desc(ordering)), group = Country ,)) +
  geom_tile(aes(y = Ref.per.pop / 2, height = Ref.per.pop, fill = Region.Name), width = 0.9) +
  geom_text(aes(label = Country), hjust = "right", colour = "black", fontface = "bold", nudge_y = -1) +
  geom_text(aes(label = scales::comma(Ref.per.pop)), hjust = "left", nudge_y = 1, colour = "grey30") +
  coord_flip(clip = "off") +
  scale_fill_manual(name = 'Region.Name', values = c("#7fc97f", "#beaed4", "#fdc086", "#ffff99", "#386cb0")) +
  scale_x_discrete("") +
  scale_y_continuous("", labels = format_si()) +
  hrbrthemes::theme_ipsum(plot_title_size = 32, subtitle_size = 24, caption_size = 12, base_size = 22) +
  theme(panel.grid.major.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        legend.position = c(0.8, 0.2),
        legend.title = element_blank(),
        plot.margin = margin(1,1,1,3,"cm"),
        axis.text.y = element_blank()) +
  # gganimate code to transition by year:
  transition_time(Year) +
  ease_aes('cubic-in-out') +
  labs(title = 'Top 10 Refugee Hosting Countries',
       subtitle = 'Ratio 1 in Refugee / 100 in Population in {round(frame_time,0)} ',
       caption = 'Source: UNHCR http://popstats.unhcr.org
Visualisation:  @edouard_lgp')

animate(p, nframes = 750, fps = 25, end_pause = 50, width = 1200, height = 900)

#anim_save("bar_chart_race_refugee_weighted.gif", animate(p, nframes = 750, fps = 20, end_pause = 60, width = 1000))

```




# Race Weighted by GDP

```{r preparedata3, message=FALSE, warning=FALSE, echo = FALSE}
rank_data.gdp <- rank_data1 %>%
    filter( !(is.na(Ref.per.gdp))) %>%
  group_by(Year, Population.type) %>%
  mutate(ordering = rank(-Ref.per.gdp)) %>%
  ungroup() %>%
  select(Country, Year, Ref.per.gdp, ordering, Value2)
rank_data.gdp <- rank_data.gdp[rank_data.gdp$ordering <= 10, ]
```

```{r chart2017facet3, message=FALSE, warning=FALSE, echo = FALSE}
ggplot(rank_data.gdp[ rank_data.gdp$Year == 2017, ]) +
  geom_bar(aes(y = Ref.per.gdp,  x =   reorder(ordering, desc(ordering)),
               group = Country ,color = Country, fill = Country), alpha = 0.75, stat = "identity") +
  
  geom_label(aes(y = 0 , x =  reorder(ordering, desc(ordering)), label = Country),
             hjust = 0,
             vjust = 0.5,
             fill = NA,
             label.size = NA,
             family = "Helvetica",
             size = 6) +
  
  #facet_wrap( ~ Population.type) +
  facet_grid(. ~ Population.type) +
  coord_flip(clip = "off", expand = FALSE) +
  scale_color_viridis_d(option = "plasma" ) +
  scale_fill_viridis_d(option = "plasma") +
  scale_y_continuous(labels = format_si()) +
  theme_minimal(14, "Avenir") +
  
  guides(color = F, fill = F) +
  labs(title =  "Top 10 Countries Refugee per GGP",
       subtitle = 'Year 2017',
       y = "Refugee to GDP Ratio",
       x = "",
       caption =  "Source: UNHCR Population Statistics -http://popstats.unhcr.org - World Bank ") +
  theme(plot.title = element_text(hjust = 1, size = 22),
        axis.ticks.y = element_blank(),
        axis.text.y  = element_blank(), 
        panel.background  = element_blank(), 
        panel.grid = element_blank(),
        plot.background = element_blank(),
        legend.position = "bottom",
        panel.grid.major.x = element_line(color = "#cbcbcb"), 
        panel.grid.major.y = element_blank()) 

```



```{r barchartrace4 , message=FALSE, warning=FALSE, echo = FALSE}
p <- ggplot(rank_data.gdp[rank_data.gdp$Population.type %in% c("Refugees (incl. refugee-like situations)"), ]) +
  geom_bar(aes(y = Ref.per.gdp,  x =   reorder(ordering, desc(ordering)),
               group = Country ,color = Country, fill = Country), alpha = 0.75, stat = "identity") +
  
  geom_label(aes(y = 0 , x =  reorder(ordering, desc(ordering)), label = Country),
             hjust = 0,
             vjust = 0.5,
             fill = NA,
             label.size = NA,
             family = "Helvetica",
             size = 6) +
  facet_grid(. ~ Population.type) +
  coord_flip(clip = "off", expand = FALSE) +
  scale_color_viridis_d(option = "plasma" ) +
  scale_fill_viridis_d(option = "plasma") +
  scale_y_continuous(labels = format_si()) +
  theme_minimal(14, "Avenir") +
  #facet_wrap( ~ Population.type) +
  #facet_grid(. ~ Population.type) +
  
  ## get the animation per year... #, nframes = 250, fps = 10, end_pause = 20
  transition_time(Year ) +
 # ease_aes('cubic-in-out') + 
 # enter_fade() +
 # exit_fade() +
  #view_follow(fixed_y = TRUE) +
  
  guides(color = F, fill = F) +
  labs(title =  "Top 10 IDPs Countries, 1993 - 2017",
       subtitle = 'Year {frame_time}',
       y = "Population Size",
       x = "",
       caption =  "Source: UNHCR Population Statistics -http://popstats.unhcr.org ") +
  theme(plot.title = element_text(hjust = 1, size = 22),
        axis.ticks.y = element_blank(),
        axis.text.y  = element_blank(), 
        panel.background  = element_blank(), 
        panel.grid = element_blank(),
        plot.background = element_blank(),
        legend.position = "bottom",
        panel.grid.major.x = element_line(color = "#cbcbcb"), 
        panel.grid.major.y = element_blank()) 

#animate(p, nframes = 250, fps = 4, end_pause = 20, width = 600)
animate(p, fps = 1, nframes = 250, end_pause = 20, width = 600)

#anim_save("bar_chart_race_idp.gif", animate(p, fps = 1, nframes = 250, end_pause = 20, width = 600))

```

# Race Weighted by GDP/Capita

```{r preparedata4, message=FALSE, warning=FALSE, echo = FALSE}
rank_data.cap <- rank_data1 %>%
    filter( !(is.na(Ref.per.gdp.cap))) %>%
  group_by(Year, Population.type) %>%
  mutate(ordering = rank(-Ref.per.gdp.cap)) %>%
  ungroup() %>%
  select(Country, Year, Ref.per.gdp.cap, ordering, Value2)
rank_data.cap <- rank_data.cap[rank_data.cap$ordering <= 10, ]

```

```{r chart2017facet4, message=FALSE, warning=FALSE, echo = FALSE}
ggplot(rank_data.cap[ rank_data.cap$Year == 2017, ]) +
  geom_bar(aes(y = Ref.per.gdp.cap,  x =   reorder(ordering, desc(ordering)),
               group = Country ,color = Country, fill = Country), alpha = 0.75, stat = "identity") +
  
  geom_label(aes(y = 0 , x =  reorder(ordering, desc(ordering)), label = Country),
             hjust = 0,
             vjust = 0.5,
             fill = NA,
             label.size = NA,
             family = "Helvetica",
             size = 6) +
  
  #facet_wrap( ~ Population.type) +
  facet_grid(. ~ Population.type) +
  coord_flip(clip = "off", expand = FALSE) +
  scale_color_viridis_d(option = "plasma" ) +
  scale_fill_viridis_d(option = "plasma") +
  scale_y_continuous(labels = format_si()) +
  theme_minimal(14, "Avenir") +
  
  guides(color = F, fill = F) +
  labs(title =  "Top 10 Countries Refugee per GDP/Cap",
       subtitle = 'Year 2017',
       y = "Refugee to GDP/Cap Ratio",
       x = "",
       caption =  "Source: UNHCR Population Statistics -http://popstats.unhcr.org - World Bank ") +
  theme(plot.title = element_text(hjust = 1, size = 22),
        axis.ticks.y = element_blank(),
        axis.text.y  = element_blank(), 
        panel.background  = element_blank(), 
        panel.grid = element_blank(),
        plot.background = element_blank(),
        legend.position = "bottom",
        panel.grid.major.x = element_line(color = "#cbcbcb"), 
        panel.grid.major.y = element_blank()) 

```



```{r barchartrace5 , message=FALSE, warning=FALSE, echo = FALSE}
p <- ggplot(rank_data.cap[rank_data.cap$Population.type %in% c("Refugees (incl. refugee-like situations)"), ]) +
  geom_bar(aes(y = Ref.per.gdp.cap,  x =   reorder(ordering, desc(ordering)),
               group = Country ,color = Country, fill = Country), alpha = 0.75, stat = "identity") +
  
  geom_label(aes(y = 0 , x =  reorder(ordering, desc(ordering)), label = Country),
             hjust = 0,
             vjust = 0.5,
             fill = NA,
             label.size = NA,
             family = "Helvetica",
             size = 6) +
  facet_grid(. ~ Population.type) +
  coord_flip(clip = "off", expand = FALSE) +
  scale_color_viridis_d(option = "plasma" ) +
  scale_fill_viridis_d(option = "plasma") +
  scale_y_continuous(labels = format_si()) +
  theme_minimal(14, "Avenir") +
  #facet_wrap( ~ Population.type) +
  #facet_grid(. ~ Population.type) +
  
  ## get the animation per year... #, nframes = 250, fps = 10, end_pause = 20
  transition_time(Year ) +
 # ease_aes('cubic-in-out') + 
 # enter_fade() +
 # exit_fade() +
  #view_follow(fixed_y = TRUE) +
  
  guides(color = F, fill = F) +
  labs(title =  "Top 10 IDPs Countries, 1993 - 2017",
       subtitle = 'Year {frame_time}',
       y = "Population Size",
       x = "",
       caption =  "Source: UNHCR Population Statistics -http://popstats.unhcr.org ") +
  theme(plot.title = element_text(hjust = 1, size = 22),
        axis.ticks.y = element_blank(),
        axis.text.y  = element_blank(), 
        panel.background  = element_blank(), 
        panel.grid = element_blank(),
        plot.background = element_blank(),
        legend.position = "bottom",
        panel.grid.major.x = element_line(color = "#cbcbcb"), 
        panel.grid.major.y = element_blank()) 

#animate(p, nframes = 250, fps = 4, end_pause = 20, width = 600)
animate(p, fps = 1, nframes = 250, end_pause = 20, width = 600)

#anim_save("bar_chart_race_idp.gif", animate(p, fps = 1, nframes = 250, end_pause = 20, width = 600))

```
